\input macros

\begindescriptions

\begindesc
\margin{This command was also described in the `Pages' chapter. The
description here now combines the two earlier descriptions.}
\bix^^{line breaks//and paragraph shape}
\easy\cts hsize {\param{dimen}}
\explain
This \minref{parameter} specifies the current ^{line length},
i.e., the usual width of lines in a paragraph
starting at the left margin.
A great many \TeX\ commands, e.g., |\centerline| (\xref{\centerline})
and |\hrule| (\xref{\hrule}), implicitly use the value of
|\hsize|.  By changing |\hsize| within a group
you can change the width of the constructs produced by such commands.

If you
set |\hsize| within a \minref{vbox} that contains text, the vbox will
have whatever width you've given to |\hsize|.
^^{vboxes//width determined by \b\tt\\hsize\e}

\PlainTeX\ sets |\hsize| to |6.5in|.

\example
{\hsize = 3.5in % Set this paragraph 3.5 inches wide.
The hedgehog was engaged in a fight with another hedgehog,
which seemed to Alice an excellent opportunity for
croqueting one of them with the other.\par}%
|
\produces
{\hsize = 3.5in
The hedgehog was engaged in a fight with another hedgehog,
which seemed to Alice an excellent opportunity for croqueting
one of them with the other.\par}%

\doruler{\8\8\8\tick\1\tick\2\tick\1\tick\3}{3.5}{in}
\nextexample
\leftline{\raggedright\vtop{\hsize = 1.5in
Here is some text that we put into a paragraph that is
an inch and a half wide.}\qquad
\vtop{\hsize = 1.5in Here is some more text that
we put into another paragraph that is an inch and a
half wide.}}
|
\produces
\leftline{\raggedright\vtop{\hsize = 1.5in
Here is some text that we put into a paragraph that is
an inch and a half wide.}\qquad
\vtop{\hsize = 1.5in Here is some more text that
we put into another paragraph that is an inch and a
half wide.}}
\endexample
\enddesc

\enddescriptions
\end