\input macros

\begindescriptions

\begindesc
\easy\ctsact ~ \xrdef{@not}
\explain
The \minref{active character} `|~|', called a ``^{tie}'',
produces a normal interword space
between two words and links those words so that
a line break will not occur between them.
You should use a tie in any context where a line break would be confusing,
e.g., before a middle initial, after an abbreviation such as ``Dr.'',
or after ``Fig.'' in ``Fig.~8''.

\example
P.D.Q.~Bach (1807--1742), the youngest and most
imitative son of Johann~S. Bach, composed the
{\sl Concerto for Horn and Hardart}.
|
\produces
\margin{The inversion of dates is deliberate---cf. Peter Schickele.}
P.D.Q.~Bach (1807--1742), the youngest and most
imitative son of Johann~S. Bach, composed the
{\sl Concerto for Horn and Hardart}.
\endexample\enddesc

\begindesc
\easy\ctspecial / \ctsxrdef{@slash}
\explain
Every character in a \TeX\ \minref{font}
has an ``^{italic correction}'' associated with it, although
the italic correction
is normally zero for a character in an unslanted (upright) font.
The italic correction specifies the extra space that's needed
when you're switching from a slanted font (not necessarily
an italic font) to an unslanted font.
The extra
space is needed because a slanted character projects into the
space that follows it, making the space look too small when the
next character is unslanted.
The metrics file for a font includes the italic correction of each
character in the font.
^^{metrics file//italic correction in}

The |\/| command
produces an ^{italic correction} for the preceding character.
You should insert an italic correction when you're switching from
a slanted font to an unslanted font,
except when the next character is a period or comma.
\example
However, {\it somebody} ate {\it something}: that's clear.

However, {\it somebody\/} ate {\it something\/}:
that's clear.
|
\produces
However, {\it somebody} ate {\it something}: that's clear.

However, {\it somebody\/} ate {\it something\/}:
that's clear.
\endexample
\enddesc

\enddescriptions
\end