\input macros

\begindescriptions

\begindesc
\cts textindent {\<argument>}
\explain
^^{indentation}
This command tells \TeX\ to start a paragraph and indent it by
|\par!-indent|,
as usual.
\TeX\ then right-justifies \<argument> within the indentation
and
follows it with an en space (half an em).
\PlainTeX\ uses this command to typeset footnotes (\xref\footnote)
^^{footnotes//using \b\tt\\textindent\e\ with}
and items in lists (see |\item|, \xref\item).

\example
\parindent = 20pt \textindent{\raise 1pt\hbox{$\bullet$}}%
You are allowed to use bullets in \TeX\ even if
you don't join the militia, and many peace-loving
typographers do so.
|
\produces
\parindent = 20pt \textindent{\raise 1pt\hbox{$\bullet$}}%
You are allowed to use bullets in \TeX\ even if
you don't join the militia, and many peace-loving
typographers do so.
\endexample\enddesc

\begindesc
\cts parindent {\param{dimen}}
\explain
This \minref{parameter} specifies the amount by which
the first line of each paragraph is to be indented. ^^{indentation}
As the example below shows, it's a bad idea to set both |\parindent|
and ^|\parskip| to zero since then the paragraph breaks are
no longer apparent.
\example
\parindent = 2em This paragraph is indented by 2 ems.
\par \parindent=0pt This paragraph is not indented at all.
\par Since we haven't reset the paragraph indentation,
this paragraph isn't indented either.
|
\produces
\parindent = 2em This paragraph is indented by 2 ems.
\par \parindent=0pt This paragraph is not indented at all.
\par Since we haven't reset the paragraph indentation,
this paragraph isn't indented either.
\endexample\enddesc

\begindesc
\cts everypar {\param{token list}}
\explain
\TeX\ performs the commands in \<token list> whenever it
enters horizontal mode, e.g., when it starts a paragraph.
By default |\everypar| is empty, but you can
take extra actions at the start of every paragraph by putting
the commands for those actions into a token list
and assigning that token list to |\everypar|.

\example
\everypar = {$\Longrightarrow$\enspace}
Now pay attention!!\par
I said, ``Pay attention!!''.\par
I'll say it again!! Pay attention!!
|
\produces
\everypar = {$\Longrightarrow$\enspace}
Now pay attention!\par
I said, ``Pay attention!''.\par
I'll say it again! Pay attention!
\endexample
\enddesc

\enddescriptions
\end