\input macros

\begindescriptions

\begindesc
\bix^^{tracing}
\bix^^{debugging}
\bix^^{diagnostic aids}
\cts show {\<token>}
\cts showthe {\<argument>}
\cts showbox {\<number>}
\cts showlists {}
\explain
These commands record information in the log of your \TeX\ run:
\ulist
\li |\show| records the meaning of \<token>.\minrefs{token}
^^{tokens//displayed by \b\tt\\show\e}
\li |\showthe| records
whatever tokens would be produced by
|\the| \<arg\-u\-ment> (see \xref \the).
\li |\showbox| records the contents of the \minref{box}
\minref{register} numbered \<num\-ber>.
The number of
leading dots in the log indicates the number of
levels of nesting of inner boxes.
\li |\showlists| records
the contents of each of the lists that \TeX\ is currently constructing.
(These lists are nested one within another.)
See \knuth{pages~88--89} for further information about interpreting
the output of |\showlists|.
\endulist
For |\show| and |\showthe|, \TeX\ also displays the information at your
^{terminal}.
For |\showbox| and |\showlists|, \TeX\ displays the information at your
terminal only
if ^|\tracingonline| (\xref \tracingonline) is greater than zero;
if ^|\tracingonline| is zero or less (the default case),
the information is not displayed.

Whenever \TeX\ encounters a |\show|-type command it
stops for interaction.  The request for interaction does \emph{not}
indicate an error, but it does give you an opportunity to ask \TeX\ to
show you something else.  If you don't want to see anything else, just
press \asciichar{return}.

You can control the amount of output produced by |\showbox| by setting
|\show!-box!-breadth| and |\show!-box!-depth| (\xref\showboxbreadth).
^^|\showboxbreadth| ^^|\showboxdepth|
These parameters respectively have default values of $5$
and $3$, which is why
just five items appear for each box described
in the log output below.  (The `|..etc.|' indicates additional items
within the boxes that aren't displayed.)
\example
\show a
\show \hbox
\show \medskip
\show &
|
\logproduces
> the letter a.
> \hbox=\hbox.
> \medskip=macro:
->\vskip \medskipamount .
> alignment tab character &.
|
\nextexample
\showthe\medskipamount
\toks27={\hbox{Joe's\quad\ Diner}}
\showthe\toks27
|
\logproduces
> 6.0pt plus 2.0pt minus 2.0pt.
> \hbox {Joe's\quad \ Diner}.
|
\nextexample
\setbox 3=\vbox{\hbox{A red dog.}\hrule A black cat.}
\showbox 3
|
\logproduces
> \box3=
\vbox(16.23332+0.0)x53.05565
.\hbox(6.94444+1.94444)x46.41675
..\tenrm A
..\glue 3.33333 plus 1.66498 minus 1.11221
..\tenrm r
..\tenrm e
..\tenrm d
..etc.
.\rule(0.4+0.0)x*
.\hbox(6.94444+0.0)x53.05565
..\tenrm A
..\glue 3.33333 plus 1.66498 minus 1.11221
..\tenrm b
..\tenrm l
..\tenrm a
..etc.
|
\endexample
\vfil\eject
\example
\vbox{A \hbox
{formula
$x \over y\showlists$}}
|
\logproduces
### math mode entered at line 3
\mathord
.\fam1 y
this will be denominator of:
\fraction, thickness = default
\\mathord
\.\fam1 x
### restricted horizontal mode entered at line 2
\tenrm f
\tenrm o
\tenrm r
\tenrm m
\kern-0.27779
\tenrm u
\tenrm l
\tenrm a
\glue 3.33333 plus 1.66666 minus 1.11111
spacefactor 1000
### horizontal mode entered at line 1
\hbox(0.0+0.0)x20.0
\tenrm A
\glue 3.33333 plus 1.66498 minus 1.11221
spacefactor 999
### internal vertical mode entered at line 1
prevdepth ignored
### vertical mode entered at line 0
prevdepth ignored
|
\endexample
\enddesc

\enddescriptions
\end