\input macros

\begindescriptions

\begindesc
\bix^^{leaders}
\easy\cts leaders {\<box or rule> \<skip command>}
\cts cleaders {\<box or rule> \<skip command>}
\cts xleaders {\<box or rule> \<skip command>}
\explain
These commands produce \minref{leaders}, i.e., they fill a
horizontal or vertical space with
copies of a pattern \seeconcept{leaders}.
The \<box> or \<rule> specifies a leader,
i.e., a single copy of the pattern,
while the \<skip command> specifies a
window to be filled with a
row or a column of the leaders.
The pattern is repeated as many times as will fit into the window.
If \<skip command> is a horizontal skip, the window
contains a row of leaders and \TeX\ must be in a horizontal mode;
if \<skip command> is a vertical skip, the window
contains a column of leaders and \TeX\ must be in a vertical mode.

The commands differ in how they arrange the repeated pattern in the space
and where they put any leftover space:

\ulist\compact
\li For |\leaders|, \TeX\ aligns a row of leaders with the left end of
the innermost \minref{box} $B$ that is to contain the result of the
|\leaders| command.
It aligns a column of leaders with the top of $B$.
Those leaders that fall entirely within the window are retained.
Any leftover space at the top and bottom of the window is left empty.
\li For |\cleaders|, the leaders are centered within the window.
\li For |\xleaders| the pattern is uniformly distributed throughout the window.
If the leftover space is $l$ and the leader is repeated  $n$ times,
\TeX\ puts space of width or height $l/(n+1)$ between adjacent leaders and
at the two ends (left and right or top and bottom) of the leaders.
\endulist

\example
\def\pattern{\hbox to 15pt{\hfil.\hfil}}
\line{Down the Rabbit-Hole {\leaders\pattern\hfil} 1}
\line{The Pool of Tears {\leaders\pattern\hfil} 9}
\line{A Caucus-Race and a Long Tale {\cleaders\pattern
\hfil} 19}
\line{Pig and Pepper {\xleaders\pattern\hfil} 27}
|
\produces
\def\pattern{\hbox to 15pt{\hfil.\hfil}}\par
\line{Down the Rabbit-Hole {\leaders\pattern\hfil} 1}
\line{The Pool of Tears {\leaders\pattern\hfil} 9}
\line{A Caucus-Race and a Long Tale {\cleaders\pattern
\hfil} 19}
\line{Pig and Pepper {\xleaders\pattern\hfil} 27}
\nextexample
\def\bulletfill{\vbox to 3ex{\vfil\hbox{$\bullet$}\vfil}}%
\def\mybox{\vbox to 1in}
\def\myrule{\hrule width 4pt}\hsize=2in
\hrule \line{%
\mybox{\myrule depth 8pt \leaders\bulletfill\vfill}
\hfil
\mybox{\myrule depth 15pt \leaders\bulletfill\vfill}
\hfil
\mybox{\myrule depth 18pt \cleaders\bulletfill\vfill}
\hfil
\mybox{\myrule depth 12pt \xleaders\bulletfill\vfill}%
}\hrule
|
\produces
\medskip
\def\bulletfill{\vbox to 3ex{\vfil\hbox{$\bullet$}\vfil}}%
\def\mybox{\vbox to 1in}\def\myrule{\hrule width 4pt}\hsize=2in
\hrule \line{%
\mybox{\myrule depth 8pt \leaders\bulletfill\vfill}
\hfil
\mybox{\myrule depth 15pt \leaders\bulletfill\vfill}
\hfil
\mybox{\myrule depth 18pt \cleaders\bulletfill\vfill}
\hfil
\mybox{\myrule depth 12pt \xleaders\bulletfill\vfill}%
}\hrule
\endexample\enddesc

\begindesc
\cts dotfill {}
\cts hrulefill {}
\explain
^^{fill}
These commands respectively fill the enclosing horizontal space with
a row of dots on the baseline and with
a horizontal line on the baseline.
It's usually a good idea to leave a space between
|\dotfill| or |\hrulefill|
and any text that precedes or follows it (see the example below).
\example
\hbox to 3in{Start {\dotfill} Finish}
\hbox to 3in{Swedish {\hrulefill} Finnish}
|
\produces
\par\hbox to 3in{Start {\dotfill} Finish}
\hbox to 3in{Swedish {\hrulefill} Finnish}
\endexample\enddesc

\begindesc
\cts leftarrowfill {}
\cts rightarrowfill {}
\explain
^^{fill}
These commands fill the enclosing horizontal space with
left-pointing or right-pointing ^{arrows}.
\example
\hbox to 3in{\vrule \rightarrowfill \ 3 in
\leftarrowfill\vrule}
|
\produces
\medskip
\hbox to 3in{\vrule \rightarrowfill \ 3 in
\leftarrowfill\vrule}
\endexample

\eix^^{leaders}
\enddesc

\enddescriptions
\end